Malaria parasites detection using mathematical morphology techniques: a review

### 2017-11-16 v1/Paper.tex
Review Andre Discussions e Conclusions

\begin{document}
	\title{Recent advances of Malaria parasites detection systems based on mathematical morphology techniques}

@@

	\begin{figure}[!t]
		\centering
		\includegraphics[width=2.75cm]{img/f3_Pfalciparum}
		\includegraphics[width=2.75cm]{img/f3_Pvivax}
		\includegraphics[width=2.75cm]{img/f3_Povale}
		\includegraphics[width=2.75cm]{img/f3_Pmalariae}
		\caption{\label{fig:malaria_types}Types of human malaria: from left to right, P. Falciparum, P. Vivax, P. Ovale, P. Malariae \cite{Khan2011}.}
	\end{figure}

	\begin{figure}[!t]
		\centering
		\includegraphics[width=2.75cm, height=2.75cm]{img/1_Pfalciparum_ring}
		\includegraphics[width=2.75cm, height=2.75cm]{img/2_Pfalciparum_trophozoite}
		\includegraphics[width=2.75cm, height=2.75cm]{img/3_Pfalciparum_gametocyte}
		\includegraphics[width=2.75cm, height=2.75cm]{img/4_Pfalciparum_schizont}
		\includegraphics[width=2.75cm, height=2.75cm]{img/5_Pmalariae_ring}
		\includegraphics[width=2.75cm, height=2.75cm]{img/6_Pmalariae_schizont}
		\includegraphics[width=2.75cm, height=2.75cm]{img/7_Povale_trophozoites}
		\includegraphics[width=2.75cm, height=2.75cm]{img/8_Pvivax_trophozoites}
		\includegraphics[width=2.75cm, height=2.75cm]{img/9_Pvivax_ring}
		\includegraphics[width=2.75cm, height=2.75cm]{img/10_Pvivax_gametocyte}
		\includegraphics[width=2.75cm, height=2.75cm]{img/11_Pvivax_ring}
		\caption{\label{fig:malaria_stages}Examples of malaria parasite stages. From top left: P.falciparum ring, trophozoite, gametocyte, schizont; P.malariae ring and schizont; P.ovale and P.vivax trophozoites; P.vivax ring and gametocyte; P.vivax ring \cite{Tek2010}}
	\end{figure}

@@

	\section{Scope of this review}
	In this paper we present a review of computer-aided methods oriented to malaria parasites detection and segmentation by mathematical morphology based techniques.
	Most of the studies were followed Di Ruberto's work \cite{DiRuberto2002}, which first proposed a system to evaluate parasitaemia in the blood. The system was able to detect the parasites by using an automatic thresholding and morphological operators. A morphological approach to cell segmentation which is more efficient than watershed algorithm was proposed. Finally, the parasites classification was still based on morphological operators. Since then many systems for computer aided diagnosis of malaria have been proposed. Most of them make use of mathematical morphology to process and analyse malaria-infected peripheral blood cells images. The scope of this paper is to review and analyze the recent works of different researchers in the area of malaria parasite recognition using computer vision which benefit from mathematical morphology.

	The rest of the paper is organised as follows. Section 2 presents a review of the considered works, according to a typical pipeline of a computer-aided image analysis process: preprocessing, segmentation, feature extraction.
	%introduces the adopted materials and methods for malaria-infected peripheral blood cells images analysis. It has been structured in preprocessing, segmentation, feature extraction, feature selection and classification subsections, consistently with a typical pipeline of a computer-aided image analysis process.
	All the considered works make use of morphological operators in at least one of the phases of image analysis. Section 3 contains an overall discussion about the methods and the conclusions are expressed in section 4.

	\section{Computer aided diagnosis of malaria by using mathematical morphology}
	This section presents a review of some of the main recent studies existing in literature regarding the analysis of malaria infected blood smears using mathematical morphology.
	A typical approach usually comprises five different image processing and analysis tasks, as follows:
	\begin{enumerate}
		\item Pre-processing.
		\item Segmentation.
		\item Feature extraction.
		\item Feature selection.
		\item Classification.
	\end{enumerate}
	Since morphological techniques have been used in the first three phases, the reviewed works have been divided into the following sub-sections: pre-processing, segmentation and feature extraction.
	Each sub-section contains description about methods that cope with malaria parasites (MP) stained components analysis, both on thin and thick blood smears, without distinction.

	%\subsection{Search of related literatures}
	Extensive search of articles has been made in PubMed and Google Scholar search engines based on the keywords: "malaria, mathematical morphology, automated malaria diagnosis" up to October  2017. The search includes papers published in English and titles and abstracts of potentially relevant studies were selected and presented from the most recent ones. Thereafter, the full texts of these studies were evaluated as per the exclusion criteria.

@@

	%\subsection{Blood smear staining techniques}
	Two main factors are generally considered if we refer to staining techniques: the type of colouration, in which Giemsa and Leishman are the most common, and the thickness of blood slide, which may be thick or thin.
	The majority of studies have been employed on thin blood smear images (over two-third of the total count) while only a few have used thick blood smear images.
	Typically, thin smears permit the identification of specific parasitic stage and quantification of malaria parasite; on the other hand, thick smears are better if the target is to perform an initial identification of malaria infection using blood pathology. Some examples are shown in fig. \ref{fig:thick_thin_film}.
	Giemsa stained blood smear is considered in most of the analysed literatures whereas Leishman stain is considered in few studies. %, as showed in table \ref{tab:summary_table}.

	It is reported that Leishman stain has bigger sensitivity for parasite detection than Giemsa \cite{Khan2014} and is superior for visualization of red and white blood cell morphology \cite{Sathpathi2014}. On the contrary, Giemsa stain highlights both malaria parasites and white blood cells and, therefore, it is an additional issue to deal with.
	Giemsa stain is much costly and also time-taking procedure than Leishman.
	Moreover, magnification of 100X by using an oil immersion objective is used for capturing microscopic images of thin blood smear for identification of specific parasites and their infected stage.

	\begin{figure}[!t]
		\centering
		\includegraphics[height=0.15\textheight]{img/f2_ThickGiemsa}
		\includegraphics[height=0.15\textheight]{img/f2_ThickLeishman}
		\includegraphics[height=0.15\textheight]{img/f2_ThinGiemsa}
		\includegraphics[height=0.15\textheight]{img/f2_ThinLeishman}
		\caption{\label{fig:thick_thin_film}Malaria infected blood smear images. From top left, clockwise: thick smear with Giemsa stain, thick smear with Leishman stain, thin smear with Leishman stain, thick smear with Giemsa stain \cite{Das2015}. Arrows in top images indicate chromatin dots while, in the bottom ones, they show the infected erythrocytes.}
	\end{figure}

	\subsection{Preprocessing}
	In image analysis field, especially when we refer to complex computer-aided pipelines, pre-processing methods are particularly used in order to improve the image data by suppressing unwanted noise or enhancing some image features for further processing.
	%Typically, pre-processing methods are classified according to the following schema:
	%\begin{enumerate}
	%	\item pixel brightness transformations;
	%	\item geometric transformations;
	%	\item pre-processing methods that use a local neighbourhood of the processed pixel;
	%	\item image restoration that requires knowledge about the entire image.
	%\end{enumerate}
	It is worth to mention pre-processing methods because they are an important step regarding image analysis field but, for what concerns the malaria-affected blood image analysis, in our review we particularly found methods which operate for illumination correction and noise filtering purposes.
	Generally speaking, digital microscopy images can be acquired in different lighting conditions, with several types of acquisition devices or from blood smears stained with various staining protocol and, consequently, the features of similar images could differ a lot.
	Different techniques for illumination correction have been suggested to reduce such variation, e.g., a lot of authors work with grayscale-converted images as an illumination correction method %\cite{Tek2010},\cite{Das2015},\cite{Tek2006}.
	On the other hand, noise filtering aims to remove the noise introduced by mishandling the slides and/or the camera settings.
	Morphological operators have been extensively used as preprocessing for image enhancement in major studies.
	Erosion and dilation operations on raw smear images allow discarding undesired patterns and help in the selection of required cells or regions of interest. Morphological operators are useful for removal of unwanted
	objects, holes filling, splitting, thinning and thinking. Different researchers during automated diagnosis of malaria used morphological operations in preprocessing phase and the most recent are listed below.

	The system proposed in \cite{Oliveira2017} is based on image processing, artificial intelligence techniques and an adapted face detection algorithm to identify Plasmodium parasites. The algorithm uses the integral image and haar-like features concepts, and weak classifiers with adaptive boosting learning. The search scope of the learning algorithm is reduced in the preprocessing step by removing the background around blood cells by means of morphological erosions both for training and for testing.
	%Morphological operations have been applied to images to enhance the geometric structures on a set of pixels. In particular, an erosion operation has been applied after a Gaussian filtering in order to locate the RBCs and restrict the search space of the algorithm.
	%The accuracy of the system was, on average, 91, meaning that for every 100 parasite-infected samples, 91 were identified correctly. The system has been developed for mobile phones and tablets and can represent a useful tool
	%in health centers in remote communities without the need for specific expertise.

	In \cite{Romero2016} an algorithm that uses morphological operations, the watershed method, the Hough transform and the clustering method of k-means to detect overlapped RBCs is presented. In the pre-processing stage white blood cells and platelets are removed before the segmentation task. During this step, some noise, the WBC cytoplasm and platelets still remain on the image. Therefore, the small objects are removed using a morphological opening and then the image is dilated with a disk-shaped structuring element.

	An edge-based segmentation of erythrocytes infected with malaria parasites using microscopic images is proposed in \cite{Somasekar2015}.  A fuzzy C-means clustering is applied to extract infected erythrocytes, which is further processed for the final segmentation. A morphological erosion is used to erase some small noises and spots before the segmentation and holes inside the infected erythrocytes are filled using a morphological hole filling operation for the final segmentation.

	The work \cite{Reni2015} proposes a new algorithm for morphological filtering of the blood images as a pre-processing tool for segmentation. Conventional morphological closing on blood images removes the unwanted components but also useful information. On the opposite the proposed method preserves the necessary information of foreground components while removing noise and artefacts.

	The paper in \cite{Kareem2012_1} illustrates a morphological approach for blood cell identification and uses the image features such as intensity, histogram, relative size and geometry for further analysis. Before the identification of blood cells, the authors  propose a novel morphological filtering based on the size of RBC for platelets and/or artifacts elimination. A dilation is performed by a concentric ring structuring element and erosion by disk shaped structuring element. The radius of the structuring element depends on the radius of the RBC, so that all the components smaller than the RBCs are removed.

	In \cite{Tek2010} a complete framework to detect and identify malaria parasites in images of Giemsa stained thin blood film specimens is proposed. Also, the system is able to identify the infecting species and life-cycle stages.
	The preprocessing step of the proposed method is applied to reduce the variations in the observed size, intensity, and colour of the cells and stained objects before the detection and classification steps. The aim is to correct the non-uniform illumination in the images. The estimation is based on a morphological closing operation using a sufficiently large structuring element. The sufficiently large size for an input image is determined automatically with respect to its average cell size computed from the area granulometry distribution.

	In the method proposed in \cite{Sheik2013} the first phase is the stained object extraction which detects candidates objects that can be infected by malaria parasites using intensity and colour. Before detecting the stained objects the method firstly extracts the foreground. Foreground image is a binary image which is produced after applying morphological hole filling on such pixels which have lower intensity value than average intensity value of green layer. After the stained objects extraction process, a series of morphological operations is also employed after every part in order to eliminate small components and complete the final stained objects.

	In \cite{Gonzalez2016} a system to determine markers for watershed segmentation based on the Radon transform and mathematical operators is proposed. In the first step of the process small irrelevant structures and part of the noise are eliminated by a morphological filter, in order to ensure the preservation of the cells edges. Image smoothing is performed by a morphological erosion-reconstruction and dilation-reconstruction filter with a disk structuring element of radius equal to 20 pixels, which is $0.274$ times smaller than the average radius of the RBCs. In this way the influences of the size and the shape of the structures can be separated in the smoothing process. At the same time the objects which are not eliminated remain unchanged. Also, a morphological closing is performed with a disc structuring element having radius smaller than half the average of the RBCs' radii, in order to connect the possible (more than one) markers that can appear on a single cell.

	%Median filter is often used for reducing impulse noise. Several studies have used it to enhance microscopic images of peripheral blood smear towards characterization of malaria followed by adaptive or local histogram equalization. Local low pass filter and local adaptive histogram equalization techniques have also been applied to enhance the pathological image quality. Das et al. \cite{Das2015} showed that geometric mean filter provides better performance towards enhancing peripheral blood smear images. Another preprocessing method like space-time manifolds-based image mosaicing has been used to obtain large field of view \cite{Zou2010} towards identifying malaria parasites, while Di Ruberto et al. \cite{Put16c} proposed an approach to overcome the problem of uneven illumination conditions in image acquisition. For this purpose they designed an illumination pattern that simulates the classic visual defects introduced by the digital microscope lenses, that
	%is the vignetting effect. Starting from the smallest radius, they applied different illumination patterns, created modifying the radius of the Gaussian curve, to the original images. Then, a similarity value measures the difference in terms of pixels between the original images and the corrupted ones.

	\subsection{Segmentation of RBCs and parasites}
	Segmentation is a key step in image analysis because it permits the identification and separation of the regions that compose an image, according to certain criteria of homogeneity and separation. Its main target is to divide the image into parts that have a strong correlation with objects or areas of the real world contained in the image.
	The commonly used segmentation methods essentially operate considering characteristics such as the brightness value, colour and reflection of the individual pixels, identifying groups of pixels that correspond to spatially connected regions. As for many problems of image processing, there is no standard solution valid in general, so different segmentation techniques can be applied, according to the characteristics of the images to process and of the objects to segment.
	Medical images segmentation is typically performed using two main strategies: the first level aims to separate whole cells or tissues from the background and the second one aims to separate the tissue structure in different regions or the cell in their components as the nucleus from the cytoplasm or intracellular parasites. The latter case is commonly used in applications in which the cell class depends on the morphological characteristics of its components.
	%The segmentation techniques can generally be divided into three main categories: Pixel Based, Edge Based and Region Based.
	%
	%%\subsection{Erythrocytes segmentation}
	%%\subsubsection{Thresholding}
	%%THRESHOLDING METHODS

	Several other authors attempted to use thresholding combined with morphological operation as segmentation method in their computer-aided systems.

	In \cite{Ross2006} the authors propose a method which provides a positive or negative diagnosis of malaria and differentiates parasites by species. The segmentation step relies on a thresholding strategy which aims to identify and segment potential parasites and erythrocytes from the image background after a six steps threshold selection. Mathematical morphology has been used for parasite size estimation, erythrocytes reconstruction and cells bigger than erythrocytes removal.
	%
	%Makkapati and Rao proposed two different methodologies. The first one \cite{Makkapati2009} is based on HSV colour space with the purpose of performing a red blood cells (RBCs) segmentation and an identification of RBCs affected from MP with a chromatin dot identification by thresholding the Saturation channel. The RBCs segmentation is achieved by dividing the entire hue range of $360^{\circ}$ into six segments and finding the dominant colour type to be the representative of the background. On the other hand, the second work \cite{Makkapati2011} is devoted to the detection of the parasite stage and identification of the species. It is based on ontological description and uses the morphology of the two basic components of a parasite, namely, chromatin dot(s) and cytoplasm. The authors state that the green component of an image offers maximum separation between the background and RBC pixels. The green component is first segmented to detect the RBCs by choosing an optimal threshold [10]. The pixels belonging to chromatin dots have higher saturation values and are segmented by thresholding the saturation component of the image. However, several artefacts may also be segmented as chromatin
	%dots, so a further evaluation is performed to establish if a chromatin dot is valid by identifying whether it is inside a RBC or not.

@@

	%The morphology of P.falciparum at gametocyte stage is the key parameter in the segmentation approach proposed in \cite{Kareem2012}. It is based on a modification of the ARR transform. The method can be used as an efficient tool for gametocyte detection for post-treatment malaria diagnosis. It also identifies the presence of any White Blood Cells (WBCs) in the image, and discards other artefacts and non infected cells. It utilizes the information based on structure, colour and geometry of the cells and does not require any segmentation or non-illumination correction techniques that are commonly used for cell detection.
	%
	%Fang et al. \cite{Fang2011} used a threshold to distinguish the unhealthy cells from the healthy cells after a malaria detection phase performed with a phase spectrum methodology. This method uses a the Quaternion Fourier Transform (QFT) to obtain the amplitude and phase spectrum for blood smear images. Then, the reconstructed image is obtained using the Inverse Quaternion Fourier transform (IQFT) on a constant amplitude spectrum and the original phase spectrum. The authors concluded that B and G channels from the RGB colour space and I channel from the HSI colour space are the most sensitive channels for MP detection.

@@

	In \cite{Ghosh2011}, authors used divergence based threshold selection in order to segment P.vivax parasites from Leishman-stained thin blood films. This method is based on Cauchy membership function \cite{DiRuberto2014} and is applied to the C channel of CMYK colour space. Morphological operators of opening and closing have been used for artefacts removal.

	Tek et al. \cite{Tek2010} addressed the problem of parasite detection and identification on thin blood film. They performed a pre-processing phase, in which morphology has been used with a granulometry-based cell size estimation, considering the peak index of the granulometry distribution as the average cell area.
	Then, two different segmentation steps followed: the foreground-background segmentation has been realized using morphological area top-hats (using the average cell area value) and morphological double thresholding; the stained pixels segmentation has been performed by modelling the stained and unstained pixel distribution with RGB space histograms and used the probability density function to determine if a pixel is stained or not. This work also addresses non-uniform illumination problem, removing it by using a pre-recorded illumination image or using a morphological closing operation.
	%
	%Malihi et al. \cite{Malihi2013} proposed a method for malaria parasite detection in Giemsa-stained blood samples. They used both binary image masks obtained with Otsu method and the edge detection images from Canny method for RBC segmentation, followed by a hole filling process.

	The authors in \cite{Arco2015} worked on thick blood films and proposed a method that uses an adaptive thresholding based scheme, which also allows an effective classification of pixels. This means that the election of whether a pixel belongs to the background or to the signal (parasites and white blood cells) is only established by the pixels around it, that is its neighbourhood. Then, morphological methods are applied to evaluate the area of connected components, labelling those belonging to parasites and counting their number. The following strategy has been used:
	\begin{enumerate}
		\item an algorithm based on run-length encoding has been used for analysing the connected components of the binary masks obtained in segmentation phase;
		\item preliminary tagging, keeping the label equivalences or connected regions with different labels in a table of local matches. Resolution of the equivalences is, then, performed;
		\item measurements on the remaining connected components: WBCs and MP region areas are checked and WBCs connected components are removed with a closing operation.
	\end{enumerate}
	%
	Also in the method proposed in \cite{Somasekar2017} the segmentation of the infected parasites is based on thresholding. The segmentation is achieved in two stages by maximizing between-class variance of an original image and consequently by an iterative threshold selection from a stage-one threshold image with suitable stopping criteria. The segmented results are post processed to improve the accuracy of the detection of malaria
	parasites by morphological operators (erosion and closing).

@@

	%% Morphology and/or granulometry in Segmentation stage:
	Di Ruberto et al. \cite{DiRuberto2002} aimed to detect the parasites by means of an automatic thresholding based on a morphological approach applied to cell image segmentation, that is more accurate than the classical watershed-based algorithm. They have applied grey scale granulometries based on opening with disk-shaped elements, flat and hemispherical. They have used a hemispherical disk-shaped structuring element to enhance the roundness and the compactness of the red cells improving the accuracy of the classical watershed algorithm, while they have used a disk-shaped flat structuring element to separate overlapping cells. These methods make use of the red blood cell structure knowledge, that is not used in existing watershed-based algorithms.

@@

@@

	In \cite{Rosado2017} Rosado et al. proposed a system using supervised classification to assess the presence of malaria parasites and determine the species and life cycle stage in Giemsa-stained thin blood smears. For the RBCs segmentation, they used an adaptive thresholding approach followed by a closing morphological operation with an elliptical structuring element.

	In \cite{Tek2010} the localisation of the parasites is achieved after a foreground and background segmentation step. Firstly, a rough foreground image using morphological area top-hats (using the average cell area value) is extracted. Then, from these rough foreground and background regions two different threshold values are determined and used in morphological double thresholding of the input grey level image to produce a refined binary foreground mask. From the foreground image the stained pixels are detected using again a thresholding approach and finally used as markers to extract the stained objects by morphological area top-hats based on the estimated average area value.

	Several authors used marker controlled watershed with morphological approach.
	Khan et al. \cite{Khan2011}, among other experimentations, used it in order to try to separate overlapping cells because, according to their statements, watershed transform can separate touching cells but it is not sufficient for overlapping cells.

>>>>

\begin{document}
\title{Malaria parasites detection using mathematical morphology techniques: a review}

@@

\begin{figure}[!t]
	\centering
	\includegraphics[width=2.75cm]{f3_Pfalciparum}
	\includegraphics[width=2.75cm]{f3_Pvivax}
	\includegraphics[width=2.75cm]{f3_Povale}
	\includegraphics[width=2.75cm]{f3_Pmalariae}
	\caption{\label{fig:malaria_types}Types of human malaria: from left to right, P. Falciparum, P. Vivax, P. Ovale, P. Malariae \cite{Khan2011}.}
\end{figure}

\begin{figure}[!t]
	\centering
	\includegraphics[width=2.75cm, height=2.75cm]{1_Pfalciparum_ring}
	\includegraphics[width=2.75cm, height=2.75cm]{2_Pfalciparum_trophozoite}
	\includegraphics[width=2.75cm, height=2.75cm]{3_Pfalciparum_gametocyte}
	\includegraphics[width=2.75cm, height=2.75cm]{4_Pfalciparum_schizont}
	\includegraphics[width=2.75cm, height=2.75cm]{5_Pmalariae_ring}
	\includegraphics[width=2.75cm, height=2.75cm]{6_Pmalariae_schizont}
	\includegraphics[width=2.75cm, height=2.75cm]{7_Povale_trophozoites}
	\includegraphics[width=2.75cm, height=2.75cm]{8_Pvivax_trophozoites}
	\includegraphics[width=2.75cm, height=2.75cm]{9_Pvivax_ring}
	\includegraphics[width=2.75cm, height=2.75cm]{10_Pvivax_gametocyte}
	\includegraphics[width=2.75cm, height=2.75cm]{11_Pvivax_ring}
	\caption{\label{fig:malaria_stages}Examples of malaria parasite stages. From top left: P.falciparum ring, trophozoite, gametocyte, schizont; P.malariae ring and schizont; P.ovale and P.vivax trophozoites; P.vivax ring and gametocyte; P.vivax ring \cite{Tek2010}}
\end{figure}

@@

In this paper we present a review of computer-aided methods oriented to malaria parasites detection and segmentation by mathematical morphology based techniques.
The rest of the paper is organised as follows. Section 2 introduces the adopted materials and methods for malaria-infected peripheral blood cells images analysis. It has been structured in preprocessing, segmentation, feature extraction, feature selection and classification subsections, consistently with a typical pipeline of a computer-aided image analysis process. All the considered works make use of morphological operators in at least one of the phases of image analysis. Section 3 contains an overall discussion about the methods and the conclusions are expressed in section 4.

\section{Materials and methods}
This section critically reviews some of the main studies existing in literature regarding the analysis of malaria infected blood smears using mathematical morphology. Since typical approaches usually comprise five different image processing and analysis tasks, the reviewed works have been divided into the following sub-sections: 
\begin{enumerate}
	\item Pre-processing.
	\item Segmentation.
	\item Feature extraction.
	\item Feature selection.
	\item Classification.
\end{enumerate}
Each sub-section contains description about methods that cope with malaria parasites (MP) stained components analysis, both on thin and thick blood smears, without distinction.

%\subsection{Search of related literatures}
Extensive search of articles has been made in PubMed and Google Scholar search engines based on the keywords: "malaria, mathematical morphology, automated malaria diagnosis" up to January 2017. The search includes papers published in English and titles and abstracts of potentially relevant studies were selected. Thereafter, the full texts of these studies were evaluated as per the exclusion criteria.

@@

%\subsection{Blood smear staining techniques}
Two main factors are generally considered if we refer to staining techniques: the type of colouration, in which Giemsa and Leishman are the most common, and the thickness of blood slide, which may be thick or thin.
The majority of studies have been employed on thin blood smear images (over two-third of the total count) while only a few have used thick blood smear images.
Typically, thin smears permit the identification of specific parasitic stage and quantification of malaria parasite; on the other hand, thick smears are better if the target is to perform an initial identification of malaria infection using blood pathology. Some examples are shown in fig. \ref{fig:thick_thin_film}.
Giemsa stained blood smear is considered in most of the analysed literatures whereas Leishman stain is considered in few studies, as showed in table \ref{tab:summary_table}.

It is reported that Leishman stain has bigger sensitivity for parasite detection than Giemsa \cite{Khan2014} and is superior for visualization of red and white blood cell morphology. \cite{Sathpathi2014}. On the contrary, Giemsa stain highlights both malaria parasites and white blood cells and, therefore, it is an additional issue to deal with. 
Giemsa stain is much costly and also time-taking procedure than Leishman.
Moreover, magnification of 100X by using an oil immersion objective is used for capturing microscopic images of thin blood smear for identification of specific parasites and their infected stage.

\begin{figure}[!t]
	\centering
	\includegraphics[height=0.15\textheight]{f2_ThickGiemsa}
	\includegraphics[height=0.15\textheight]{f2_ThickLeishman}
	\includegraphics[height=0.15\textheight]{f2_ThinGiemsa}
	\includegraphics[height=0.15\textheight]{f2_ThinLeishman}
	\caption{\label{fig:thick_thin_film}Malaria infected blood smear images. From top left, clockwise: thick smear with Giemsa stain, thick smear with Leishman stain, thin smear with Leishman stain, thick smear with Giemsa stain \cite{Das2015}. Arrows in top images indicate chromatin dots while, in the bottom ones, they show the infected erythrocytes.}
\end{figure}

\subsection{Preprocessing}
In image analysis field, especially when we refer to complex computer-aided pipelines, pre-processing methods are particularly used in order to improve the image data by suppressing unwanted noise or enhancing some image features for further processing.
Typically, pre-processing methods are classified according to the following schema:
\begin{enumerate}
	\item pixel brightness transformations;
	\item geometric transformations;
	\item pre-processing methods that use a local neighbourhood of the processed pixel;
	\item image restoration that requires knowledge about the entire image.
\end{enumerate}
It is worth to mention pre-processing methods because they are an important step regarding image analysis field but, for what concerns the malaria-affected blood image analysis, in our review we particularly found methods which operate for illumination correction and noise filtering purposes.
Generally speaking, digital microscopy images can be acquired in different lighting conditions, with several types of acquisition devices or from blood smears stained with various staining protocol and, consequently, the features of similar images could differ a lot.
Different techniques for illumination correction have been suggested to reduce such variation, e.g., a lot of authors work with grayscale-converted images as an illumination correction method \cite{Tek2010},\cite{Das2015},\cite{Tek2006}.
On the other hand, noise filtering aims to remove the noise introduced by mishandling the slides and/or the camera settings.
Median filter is often used for reducing impulse noise. Several studies have used it to enhance microscopic images of peripheral blood smear towards characterization of malaria followed by adaptive or local histogram equalization. Local low pass filter and local adaptive histogram equalization techniques have also been applied to enhance the pathological image quality. Das et al. \cite{Das2015} showed that geometric mean filter provides better performance towards enhancing peripheral blood smear images. Another preprocessing method like space-time manifolds-based image mosaicing has been used to obtain large field of view \cite{Zou2010} towards identifying malaria parasites, while Di Ruberto et al. \cite{Put16c} proposed an approach to overcome the problem of uneven illumination conditions in image acquisition. For this purpose they designed an illumination pattern that simulates the classic visual defects introduced by the digital microscope lenses, that
is the vignetting effect. Starting from the smallest radius, they applied different illumination patterns, created modifying the radius of the Gaussian curve, to the original images. Then, a similarity value measures the difference in terms of pixels between the original images and the corrupted ones.

\subsection{Segmentation}
Segmentation is a key step in image analysis because it permits the identification and separation of the regions that compose an image, according to certain criteria of homogeneity and separation. Its main target is to divide the image into parts that have a strong correlation with objects or areas of the real world contained in the image.
The commonly used segmentation methods essentially operate considering characteristics such as the brightness value, colour and reflection of the individual pixels, identifying groups of pixels that correspond to spatially connected regions. As for many problems of image processing, there is no standard solution valid in general, so different segmentation techniques can be applied, according to the characteristics of the images to process and of the objects to segment. 
Medical images segmentation is typically performed using two main strategies: the first level aims to separate whole cells or tissues from the background and the second one aims to separate the tissue structure in different regions or the cell in their components as the nucleus from the cytoplasm or intracellular parasites. The latter case is commonly used in applications in which the cell class depends on the morphological characteristics of its components. 
The segmentation techniques can generally be divided into three main categories: Pixel Based, Edge Based and Region Based.

%\subsection{Erythrocytes segmentation}
%\subsubsection{Thresholding}
%THRESHOLDING METHODS
Several other authors attempted to use thresholding as segmentation method in their computer-aided systems.
In \cite{Ross2006}, the authors propose a method which provides a positive or negative diagnosis of malaria and differentiates parasites by species. The segmentation step relies on a thresholding strategy which aims to identify and segment potential parasites and erythrocytes from the image background after a six steps threshold selection. Mathematical morphology has been used for parasite size estimation, erythrocytes reconstruction and cells bigger than erythrocytes removal.

Makkapati and Rao proposed two different methodologies. The first one \cite{Makkapati2009} is based on HSV colour space with the purpose of performing a red blood cells (RBCs) segmentation and an identification of RBCs affected from MP with a chromatin dot identification by thresholding the Saturation channel. The RBCs segmentation is achieved by dividing the entire hue range of $360^{\circ}$ into six segments and finding the dominant colour type to be the representative of the background. On the other hand, the second work \cite{Makkapati2011} is devoted to the detection of the parasite stage and identification of the species. It is based on ontological description and uses the morphology of the two basic components of a parasite, namely, chromatin dot(s) and cytoplasm. The authors state that the green component of an image offers maximum separation between the background and RBC pixels. The green component is first segmented to detect the RBCs by choosing an optimal threshold [10]. The pixels belonging to chromatin dots have higher saturation values and are segmented by thresholding the saturation component of the image. However, several artefacts may also be segmented as chromatin
dots, so a further evaluation is performed to establish if a chromatin dot is valid by identifying whether it is inside a RBC or not.

@@

The morphology of P.falciparum at gametocyte stage is the key parameter in the segmentation approach proposed in \cite{Kareem2012}. It is based on a modification of the ARR transform. The method can be used as an efficient tool for gametocyte detection for post-treatment malaria diagnosis. It also identifies the presence of any White Blood Cells (WBCs) in the image, and discards other artefacts and non infected cells. It utilizes the information based on structure, colour and geometry of the cells and does not require any segmentation or non-illumination correction techniques that are commonly used for cell detection.

Fang et al. \cite{Fang2011} used a threshold to distinguish the unhealthy cells from the healthy cells after a malaria detection phase performed with a phase spectrum methodology. This method uses a the Quaternion Fourier Transform (QFT) to obtain the amplitude and phase spectrum for blood smear images. Then, the reconstructed image is obtained using the Inverse Quaternion Fourier transform (IQFT) on a constant amplitude spectrum and the original phase spectrum. The authors concluded that B and G channels from the RGB colour space and I channel from the HSI colour space are the most sensitive channels for MP detection.

@@

In \cite{Ghosh2011}, authors used divergence based threshold selection in order to segment P.vivax parasites from Leishman-stained thin blood films. This method is based on Cauchy membership function \cite{Put14a} and is applied to the C channel of CMYK colour space. Morphological operators of opening and closing have been used for artefacts removal.

Tek et al. \cite{Tek2010} addressed the problem of parasite detection and identification on thin blood film. They performed a pre-processing phase, in which morphology has been used with a granulometry-based cell size estimation, considering the peak index of the granulometry distribution as the average cell area.
Then, two different segmentation steps followed: the foreground-background segmentation has been realized using morphological area top-hats (using the average cell area value) and morphological double thresholding; the stained pixels segmentation has been performed by modelling the stained and unstained pixel distribution with RGB space histograms and used the probability density function to determine if a pixel is stained or not. This work also addresses non-uniform illumination problem, removing it by using a pre-recorded illumination image or using a morphological closing operation.

Malihi et al. \cite{Malihi2013} proposed a method for malaria parasite detection in Giemsa-stained blood samples. They used both binary image masks obtained with Otsu method and the edge detection images from Canny method for RBC segmentation, followed by a hole filling process.

Authors in \cite{Arco2015} worked on thick blood films and proposed a method that uses an adaptive thresholding based scheme, which also allows an effective classification of pixels. This means that the election of whether a pixel belongs to the background or to the signal (parasites and white blood cells) is only established by the pixels around it, that is its neighbourhood. Then, morphological methods are applied to evaluate the area of connected components, labelling those belonging to parasites and counting their number. The following strategy has been used:
\begin{enumerate}
	\item an algorithm based on run-length encoding has been used for analysing the connected components of the binary masks obtained in segmentation phase;
	\item preliminary tagging, keeping the label equivalences or connected regions with different labels in a table of local matches. Resolution of the equivalences is, then, performed;
	\item measurements on the remaining connected components: WBCs and MP region areas are checked and WBCs connected components are removed with a closing operation.
\end{enumerate}

@@

% Morphology and/or granulometry in Segmentation stage:
Di Ruberto et al. \cite{DiRuberto2002} aimed to detect the parasites by means of an automatic thresholding based on a morphological approach applied to cell image segmentation, that is more accurate than the classical watershed-based algorithm. They have applied grey scale granulometries based on opening with disk-shaped elements, flat and hemispherical. They have used a hemispherical disk-shaped structuring element to enhance the roundness and the compactness of the red cells improving the accuracy of the classical watershed algorithm, while they have used a disk-shaped flat structuring element to separate overlapping cells. These methods make use of the red blood cell structure knowledge, that is not used in existing watershed-based algorithms.

@@

Several authors used marker controlled watershed with morphological approach.
Khan et al. \cite{Khan2011}, among other experimentations, used it in order to try to separate overlapping cells because, according to their statements, watershed transform can separate touching cells but it is not sufficient for overlapping cells.
In \cite{Springl2009} red blood cell segmentation has been performed by using marker-controlled watershed transformation based on the image gradient. Markers are computed as a combination of the binary mask of the red blood cells and centres of the cells which are computed using a similar algorithm that was utilized for the evaluation of the average cell radius. The binary mask is obtained by thresholding the grey-scale image with an automatically estimated threshold using Otsu method\cite{Otsu1975}.
Das et al. \cite{Das2013}, \cite{Das2012}, \cite{Das2011}, \cite{Das2015} segmented erythrocytes as aforesaid and then morphological operators are used to eliminate unwanted cells like leukocytes and platelets. To conclude, overlapping erythrocytes are segmented by using marker controlled watershed segmentation technique.

%BOUNDARY BASED METHODS \subsubsection{Boundary based methods}
Zou et al. \cite{Zou2010} based their method on the consideration that RBCs have a circular shape. They propose an improved version of the Circle Hough Transform (CHT) for RBCs detection. The authors state that overlapping and oval cells can be detected as well as non-clumped RBCs and affirm that the green channel of the RGB colour space permits to better distinguish RBCs. However, some RBCs may not have circular shape and it could affect the detection results.

@@

%CLUSTERING METHODS \subsection{Clustering methods}: no method uses mathematical morphology applied to clustering.

%CLASSIFIER-BASED METHODS \subsection{Classifier-based methods}
Tek et al. \cite{Tek2006} proposed a classifier-based method, for the segmentation stage, which relies on a Bayesian pixel classifier to distinguish among stained and non-stained pixels. In particular, they used a non-parametric method based on histograms in order to produce the probability density functions of stained and non-stained classes. Stained pixels can belong to other components such as WBCs, platelets or artefacts, in addition to the parasites and so the detection procedure requires a further classification to distinguish among parasite and non-parasite pixels. However, the stained pixels have to be represented as connected sets, representing stained objects, to extract features for the classifier. Furthermore, top-hat extraction and infinite reconstruction were applied to find the regions that include the objects.
